---
title: "p8105_hw2_cm3928"
author: "Clement Mugenzi"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(readr)
library(haven)
library(skimr)
```

# Question 1

## First Part

In the chunk below, I will load the dataset, clean names, and get rid of unwanted rows.

```{r, include=TRUE}
trash_df = 
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  filter(dumpster >= 1) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0),
    sports_balls = as.integer(sports_balls)
  )
view(trash_df)
```

The chunk below converts the variable sports_balls from decimal to integer.

## Second Part

Let us first load and clean precipitation data from year 2018 as well as add a new variable year.

```{r, include=TRUE}
prec_2018 =
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             range = "A2:B14", sheet = 5) %>% 
  mutate(year = 2018) %>%
  janitor::clean_names() %>% 
  drop_na(total)
view(prec_2018)
```

Next, I will load and clean precipitation data from 2017 as well as create a new variable year.

```{r, include=TRUE}
prec_2017 = 
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             range = "A2:B14", sheet = 6) %>% 
  mutate(year = 2017) %>% 
  janitor::clean_names()
view(prec_2017)
```


Finally, I will combine both the 2018 and 2017 precipitation dataframes and convert month to character variable.

```{r, include=TRUE}
prec_df = 
  left_join(prec_2018, prec_2017, by = "month") %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month])
view(prec_df)
```




# Question 2 

**First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.**

```{r, include=TRUE}
pols_df = 
  read_excel("fivethirtyeight_datasets/pols-month.xlsx") %>% 
  separate(col = mon, into = c("year", "month", "day")) %>% 
   mutate(
    month = as.numeric(month),
    month = month.abb[month],
    president = recode(prez_gop, `0` = "dem", `1` = "gop")) %>% 
  select(-c(prez_gop, prez_dem, day)) %>%
view(pols_df)
```


**Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.**

```{r, include=TRUE}
snp_df1 = 
  read_csv("fivethirtyeight_datasets/snp.csv") %>%   
  separate(col = date, into = c("day", "month", "year")) %>% 
  mutate(
    month = as.numeric(month),
    month = month.abb[month]) %>% 
  select(-day)
snp_df = snp_df1[,c(2,1,3)]
view(snp_df)
```


**Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.**

```{r, include=TRUE}
jobless_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "day") %>% 
  select(-day)
# how do you represent day as integer. not possible so far
view(jobless_df)
```

**Join the datasets by merging snp into pols, and merging unemployment into the result.**

Let us first join pols_df and snp_df.

```{r, include=TRUE}
df = 
  left_join(snp_df, pols_df, by = c("month", "year")) %>% 
              mutate(
                year = as.numeric(year)
)
view(df)
```

Next, we will join the new dataframe created in the above chunk with the jobless/unemployment dataframe.

```{r}
master_df = 
  left_join(df, jobless_df, by = c("month", "year"))
view(master_df)
```


# Question 3

```{r}
# Loading and cleaning the dataset as well as getting rid of rows which appears twice.
popnames_df = 
  read_csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct() %>%
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name)) %>% 
  mutate(
    ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander"),
    ethnicity = recode(ethnicity, "black non hisp" = "black non hispanic"),
    ethnicity = recode(ethnicity, "white non hisp" = "white non hispanic")
  )
view(popnames_df)
```


Let's create a table using **tibble** to showcase the frequency of the name **Olivia** over time among female subjects.

```{r, include=TRUE}
olivia_as_name = 
  popnames_df %>% 
  filter(childs_first_name == "olivia") %>% 
  select(-count, -gender) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )
view(olivia_as_name)
```

Next, I will produce a nice, readable tibble representing popular male's name by ethnicity and year of birth.

```{r}
pop_male_name = 
  popnames_df %>% 
  filter(rank == "1", gender == "male") %>% 
  select(-gender, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )
view(pop_male_name)
```

Finally, I will produce a scatterplot for white non hispanic male born in 2016, showing the number of children with a name against the rank in popularity of that name.

```{r}
scatter_male = 
  popnames_df %>% 
  filter(gender == "male", 
         ethnicity == "white non hispanic", 
         year_of_birth == "2016") %>% 
  ggplot(aes(x = rank, y = count)) +
           geom_point()
scatter_male
```





























