---
title: "p8105_hw2_cm3928"
author: "Clement Mugenzi"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(readr)
library(haven)
library(skimr)
```

# Question 1

## First Part

In the chunk below, I will load the dataset, clean names, and get rid of unwanted rows.

```{r, include=TRUE}
trash_df = 
  read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx",
             range = "A2:N338") %>% 
  janitor::clean_names() %>% 
  filter(dumpster >= 1)
view(trash_df)
```

The chunk below converts the variable sports_balls from decimal to integer.

```{r, include=TRUE}
sports_b = round(pull(trash_df, sports_balls), digits = 2)
sports_int = as.integer(sports_b)
```

## Second Part

Let us first load and clean precipitation data from year 2018 as well as add a new variable year.

```{r}
prec_2018_pre =
  read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx",
             range = "A2:B9", sheet = 3) %>% 
  mutate(year = 2018) %>%
  janitor::clean_names()
prec_2018 = prec_2018_pre[,c(3,1,2)]
view(prec_2018)
```

Next, I will load and clean precipitation data from 2017 as well as create a new variable year.

```{r}
prec_2017_pre = 
  read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx",
             range = "A2:B14", sheet = 4) %>% 
  mutate(year = 2017) %>% 
  janitor::clean_names()
prec_2017 = prec_2017_pre[,c(3,1,2)]
view(prec_2017)
```






















