---
title: "p8105_hw2_cm3928"
author: "Clement Mugenzi"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(readr)
library(haven)
library(skimr)
```

# Question 1

## First Part

In the chunk below, I will load the dataset, clean names, and get rid of unwanted rows.

```{r, include=TRUE}
trash_df = 
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  filter(dumpster >= 1) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0),
    sports_balls = as.integer(sports_balls)
  )
view(trash_df)
```


```{r}
kable(trash_df, markdown, digits = 1)
```

## Second Part

Let us first load and clean precipitation data from year 2018 as well as add a new variable year.

```{r, include=TRUE}
prec_2018 =
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             range = "A2:B14", sheet = 5) %>%
  mutate(year = 2018) %>% 
  janitor::clean_names()
```

```{r}
kable(prec_2018, markdown, digits = 1)
```

Next, I will load and clean precipitation data from 2017 as well as create a new variable year.

```{r, include=TRUE}
prec_2017 = 
  read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             range = "A2:B14", sheet = 6) %>% 
  mutate(year = 2017) %>% 
  janitor::clean_names()
```

```{r}
kable(prec_2017, markdown, digits = 1)
```


Finally, I will combine both the 2018 and 2017 precipitation dataframes and convert month to character variable.

```{r, include=TRUE}
prec_df = 
  left_join(prec_2017, prec_2018, by = "month") %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month])
```


```{r}
kable(prec_df, markdown, digits = 1)
```

The Mr. Trash Wheel is a dataset comprised of **`r ncol(trash_df)`** variables and **`r nrow(trash_df)`** observations. And the key variables are **dumpster**, **month**, **year** and **sports balls**.

Precipitation dataframes for years 2018 and 2017, both have **`r ncol(prec_2018)`** variables and **`r nrow(prec_2017)`** observations. Lastly, when both precipitation dataframes are combined, the resulting dataframe is comprised of **`r ncol(prec_df)`** variables and **`r nrow(prec_df)`** observations. Key variables for the combined dataframe are **month** and **year**.

The total precipitation for year 2018 was **`r sum(pull(prec_2018, total))`**.  

```{r}
med = 
  trash_df %>% 
  filter(year == "2017") %>% 
  select(year, sports_balls)
```

So, the median number of sports balls for year 2017 was **`r median(med$sports_balls)`**.


# Question 2 

The Chunk below will load the dataset 'pols_month', clean it, and break up the variable 'mon' into variables year, month, and day. It will also replace month number with month name, and finally, create a president variable with values gop and dem.

```{r, include=TRUE}
pols_df = 
  read_excel("fivethirtyeight_datasets/pols-month.xlsx") %>% 
  separate(col = mon, into = c("year", "month", "day")) %>% 
   mutate(
    month = as.numeric(month),
    month = month.name[month],
    president = recode(prez_gop, `0` = "dem", `1` = "gop")) %>% 
  select(-c(prez_gop, prez_dem, day))
```


```{r}
kable(pols_df, markdown, digits = 1)
```

The following chunk will load, clean the dataset 'snp', and variables will be arranged in a way to make it easier for me to join datasets.

```{r, include=TRUE}
snp_df1 = 
  read_csv("fivethirtyeight_datasets/snp.csv") %>%   
  separate(col = date, into = c("day", "month", "year")) %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month]) %>% 
  select(-day)
snp_df = snp_df1[,c(2,1,3)]
```

```{r}
kable(snp_df, markdown, digits = 1)
```

The chunk below will load and clean the unemployment dataset, and finally, tidy it with a goal of being able to merge it with the two datasets above.

```{r, include=TRUE}
jobless_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "day") %>% 
  select(-day)
view(jobless_df)
```

Let us first join pols_df and snp_df.

```{r, include=TRUE}
df = 
  left_join(snp_df, pols_df, by = c("month", "year")) %>% 
  mutate(
    year = as.numeric(year)
)
view(df)
```

Next, we will join the new dataframe created in the above chunk with the jobless/unemployment dataframe.

```{r}
master_df = 
  left_join(df, jobless_df, by = c("month", "year"))
view(master_df)
```

```{r}
kable(master_df, markdown)
```


For the pols-month (later named pols_df), the number of variables and observations did not change when the original dataset is compared to the final (9 variables and 822 observations). Range of years is **`r range(pull(pols_df, year))`**. Names of key variables are month, year, and president.

For the snp dataframe, I started with 2 columns and 787 observations and ended up with a total of **`r ncol(snp_df)`** variables and **`r nrow(snp_df)`** observations. Range of years is **`r range(pull(snp_df, year))`**. And finally, key variables are year and month.

For the unemployment (jobless_df) dataframe, I started with 13 variables and 68 observations and ended up with **`r ncol(jobless_df)`** variables and **`r nrow(jobless_df)`** observations. Range of years is **`r range(pull(jobless_df, year))`** and key variables are year and month.

For the master dataframe (master_df), the resulting variables are **`r ncol(master_df)`** and observations are **`r nrow(master_df)`**. Range of years are **`r range(pull(master_df, year))`**. Key variables being month, year, and president.

# Question 3

```{r}
# Loading and cleaning the dataset as well as getting rid of rows which appears twice.
popnames_df = 
  read_csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct() %>%
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name)) %>% 
  mutate(
    ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander"),
    ethnicity = recode(ethnicity, "black non hisp" = "black non hispanic"),
    ethnicity = recode(ethnicity, "white non hisp" = "white non hispanic")
  )
```


Let's create a table using **tibble** to showcase the frequency of the name **Olivia** over time among female subjects.

```{r, include=TRUE}
olivia_as_name = 
  popnames_df %>% 
  filter(childs_first_name == "olivia") %>% 
  select(-count, -gender) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )
```

```{r}
kable(olivia_as_name, markdown, digits = 1)
```

Next, I will produce a nice, readable tibble representing popular male's name by ethnicity and year of birth.

```{r}
pop_male_name = 
  popnames_df %>% 
  filter(rank == "1", gender == "male") %>% 
  select(-gender, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )
```

```{r}
kable(pop_male_name, markdown, digits = 1)
```

Finally, I will produce a scatterplot for white non hispanic male born in 2016, showing the number of children with a name against the rank in popularity of that name.

```{r}
scatter_male = 
  popnames_df %>% 
  filter(gender == "male", 
         ethnicity == "white non hispanic", 
         year_of_birth == "2016") %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point() +
  labs(
    title = "Scatter Plot for non Hispanic Male Born in 2016",
    x = "Rank in Popularity",
    y = "Number of Children with a given name"
  )
scatter_male
```




























